---
title: "Data visualization overview"
subtitle: "Lecture 4"
date: last-modified
date-format: long
format: 
  revealjs:
    footer: "[🔗 renr213.github.io](https://renr213.github.io/)"
---


```{r}
#| echo: false

library(countdown)
library(waffle)
library(ggthemes)
library(tidyverse)

ggplot2::theme_set(ggplot2::theme_gray(base_size = 16))
```

## Packages

```{r}
library(palmerpenguins)
library(tidyverse)
library(ggthemes)
```

## Bivariate analysis {.smaller}

:::::: panel-tabset
## Plot

::::: columns
::: {.column width="50%"}
```{r}
#| label: penguins-box
#| echo: false
#| fig-width: 5
#| fig-asp: 0.8

# Side-by-side box plots
ggplot(penguins, aes(x = body_mass_g, y = species, fill = species)) +
  geom_boxplot(alpha = 0.5, show.legend = FALSE) +
  scale_fill_colorblind() +
  labs(
    x = "Body mass (grams)", y = "Species",
    title = "Side-by-side box plots"
  )
```
:::

::: {.column width="50%"}
```{r}
#| label: penguins-density
#| echo: false
#| fig-width: 5
#| fig-asp: 0.8

# Density plots
ggplot(penguins, aes(x = body_mass_g, fill = species)) +
  geom_density(alpha = 0.5) +
  theme(legend.position = "bottom") +
  scale_fill_colorblind() +
  labs(
    x = "Body mass (grams)", y = "Density",
    fill = "Species", title = "Density plots"
  )
```
:::
:::::

## Code

```{r}
#| ref.label: penguins-box
#| fig-show: hide
```

```{r}
#| ref.label: penguins-density
#| fig-show: hide
```
::::::

## Violin plots

::::: columns
::: {.column width="50%"}
```{r}
#| label: penguins-violin-1
#| fig-show: hide
#| code-line-numbers: "|4|5|8"

ggplot(
  penguins,
  aes(
    x = species,
    y = body_mass_g
    )
  ) +
  geom_violin()
```
:::

::: {.column width="50%"}
```{r}
#| ref.label: penguins-violin-1
#| echo: false
#| fig-width: 5
#| fig-asp: 0.8
```
:::
:::::

## Multiple geoms

::::: columns
::: {.column width="50%"}
```{r}
#| label: penguins-violin-2
#| fig-show: hide
#| code-line-numbers: "9"

ggplot(
  penguins,
  aes(
    x = species,
    y = body_mass_g
    )
  ) +
  geom_violin() +
  geom_point()
```
:::

::: {.column width="50%"}
```{r}
#| ref.label: penguins-violin-2
#| echo: false
#| fig-width: 5
#| fig-asp: 0.8
```
:::
:::::

## Multiple geoms

::::: columns
::: {.column width="50%"}
```{r}
#| label: penguins-violin-3
#| fig-show: hide
#| code-line-numbers: "9"

ggplot(
  penguins,
  aes(
    x = species,
    y = body_mass_g
    )
  ) +
  geom_violin() +
  geom_jitter()
```
:::

::: {.column width="50%"}
```{r}
#| ref.label: penguins-violin-3
#| echo: false
#| fig-width: 5
#| fig-asp: 0.8
```
:::
:::::

## Multiple geoms + aesthetics

::::: columns
::: {.column width="50%"}
```{r}
#| label: penguins-violin-4
#| fig-show: hide
#| code-line-numbers: "5-6"

ggplot(
  penguins,
  aes(
    x = species,
    y = body_mass_g,
    color = species
    )
  ) +
  geom_violin() +
  geom_jitter()
```
:::

::: {.column width="50%"}
```{r}
#| ref.label: penguins-violin-4
#| echo: false
#| fig-width: 5
#| fig-asp: 0.8
```
:::
:::::

## Multiple geoms + aesthetics

::::: columns
::: {.column width="50%"}
```{r}
#| label: penguins-violin-5
#| fig-show: hide
#| code-line-numbers: "11-13"

ggplot(
  penguins,
  aes(
    x = species,
    y = body_mass_g,
    color = species
    )
  ) +
  geom_violin() +
  geom_jitter() +
  theme(
    legend.position = "none"
  )
```
:::

::: {.column width="50%"}
```{r}
#| ref.label: penguins-violin-5
#| echo: false
#| fig-width: 5
#| fig-asp: 0.8
```
:::
:::::

## Multiple geoms + aesthetics {.smaller}

::::: columns
::: {.column width="50%"}
```{r}
#| label: penguins-violin-6
#| fig-show: hide
#| code-line-numbers: "14"

ggplot(
  penguins,
  aes(
    x = species,
    y = body_mass_g,
    color = species
    )
  ) +
  geom_violin() +
  geom_jitter() +
  theme(
    legend.position = "none"
  ) +
  scale_color_colorblind()
```
:::

::: {.column width="50%"}
```{r}
#| ref.label: penguins-violin-6
#| echo: false
#| fig-width: 5
#| fig-asp: 0.8
```
:::
:::::

# Multivariate analysis

## Bechdel {.smaller}

Load the Bechdel test data with `read_csv()`:

```{r}
#| label: load-bechdel
#| message: false

bechdel <- read_csv("data/bechdel.csv")
```

. . .

<br>

View the column `names()` of the `bechdel` data:

```{r}
#| label: names-bechdel

names(bechdel)
```

## ROI by test result {.smaller}

::: question
What about this plot makes it difficult to evaluate how ROI varies by Bechdel test result?
:::

```{r}
#| label: roi-test-binary-box
#| fig-width: 8
#| fig-asp: 0.4

ggplot(bechdel, aes(x = roi, y = clean_test, color = binary)) +
  geom_boxplot()
```

## Movies with high ROI {.smaller}

::: question
What are the movies with highest ROI?
:::

```{r}
bechdel |>
  filter(roi > 400) |>
  select(title, roi, budget_2013, gross_2013, year, clean_test)
```

## ROI by test result {.smaller}

::: question
**Zoom in:** What about this plot makes it difficult to evaluate how ROI varies by Bechdel test result?
:::

```{r}
#| label: roi-test-binary-box-zoom
#| fig-width: 8
#| fig-asp: 0.4
#| code-line-numbers: "|3"

ggplot(bechdel, aes(x = roi, y = clean_test, color = binary)) +
  geom_boxplot() +
  coord_cartesian(xlim = c(0, 15))
```

## Sneak preview...

<br> <br>

::: hand
of data wrangling pipelines...
:::

## Median ROI

```{r}
#| label: median-roi
#| code-line-numbers: "|2"

bechdel |>
  summarize(median_roi = median(roi, na.rm = TRUE))
```

## Median ROI by test result

```{r}
#| label: median-roi-test
#| code-line-numbers: "|2|3"

bechdel |>
  group_by(clean_test) |>
  summarize(median_roi = median(roi, na.rm = TRUE))
```

## ROI by test result – zoom in {.smaller}

::: question
What does this plot say about return-on-investment on movies that pass the Bechdel test?
:::

```{r}
#| label: roi-test-binary-box-zoom-vline
#| fig-width: 9
#| fig-asp: 0.4

ggplot(bechdel, aes(x = roi, y = clean_test, color = binary)) +
  geom_boxplot() +
  coord_cartesian(xlim = c(0, 15)) +
  geom_vline(xintercept = 4.21, linetype = "dashed")
```

# Application exercise

## `ae-03-duke-forest` {.smaller}
